---
output:
  pdf_document:
    latex_engine: xelatex
---


# Title: PH125X9 Capston Project II: Prediction of liver disease 
## by JING JING SUN
## OCT. 19, 2021

# **I.Introduction**

Prediction of the  human disease based on clinic laboratory data is possible as data science developing.  It will play an important role in public health,  estimate the risk and diagnosis of disease. 


I downloaded a India Liver Patient Dataset(ILPD)  from the University of California , Irvine
(UCI) machine learning repository(https://archive.ics.uci.edu/ml/machine-learning-databases/00225/Indian%20Liver%20Patient%20Dataset%20(ILPD).csv).

This data set is collected from North East of Andhra Pradesh, India. There are 416 liver patient records and 167 non-liver patient records.  Data set columns include Age of the patient, Gender of the patient, total Bilirubin, Direct Bilirubin, Alkaline phosphatase, Alamine Aminotransferase(ALT), Aspatate Aminotransferase(AST), Total protein, Albumin and Albumin and Globulin ratio.

The goal of this project is to create a disease prediction system use patient laboratory test result to create a disease prediction system, with the knowledge we have learned from R course series and help us get the ability to contribute to real world. 

# **II. Method/Analysis**

## **loading needed packages**
```{r}
library(tidyverse)
library(dslabs)
library(matrixStats)
library(caret)
library(lubridate)
library(dplyr)
```
## ** data clean and organization **
## **load dataset**

I go to the University of California , Irvine website, read the a India Liver Patient Dataset(ILPD).    
```{r}
liver_data <- read.csv("https://archive.ics.uci.edu/ml/machine-learning-databases/00225/Indian%20Liver%20Patient%20Dataset%20(ILPD).csv", 
                       header = FALSE)
head(liver_data)

```


## **give each column a name**
I noticed there are no column name in the data set, so I gave each column a name
```{r}
colnames(liver_data) <- c("Age", "Sex", "Tot_Bil", "Dir_Bil", "Alkphos", "Alamine", 
                          "Aspartate", "Tot_Prot", "Albumin", "A_G_Ratio", "Disease")

```

## **convert value in column disease into 0 (no disease) and 1 ( liver disease)**
In order to process data later, I use zero and one to represent no disease and with  liver disease. 
```{r}
liver_data$Disease <- as.numeric(ifelse(liver_data$Disease == 2, 0, 1))  
head(liver_data)
```
## **add column "liverdisease" as.factor "0" represent "nodisease", "1" represent "liverdisease"**
 I also add a column to as.factor zero and one to represent no disease and  liver disease.
```{r}
liver_data<- liver_data %>% mutate ( liver_data, liverdisease= as.factor (Disease))
head(liver_data)
```

## **data set exploration**
```{r}
str(liver_data)
dim(liver_data)
```
This data set have 583 row and  12 column.

## **remove those rows that have missing data**
```{r}
liver_data<-na.omit(liver_data)
dim(liver_data)
```
Now this data set have 579 row and  12 column. We removed 4 rows from the dataset.
 
 
## **know our cleaned data**
```{r}
summary(liver_data)
```

## **how many kind of liver disease in this data set**
```{r}
n_distinct(liver_data$Disease)
histogram(liver_data$Disease)
```
There are only 2 situations: no disease(0), liver disease(1).

## **how many liver disease case**
```{r}
count(liver_data$Disease == "1") 
```
## **how many no disease case**
```{r}
sum(liver_data$Disease == "0")
```
## **distribution of Age**
```{r}
liver_data%>% ggplot(aes(Age)) +
  geom_histogram( binwidth = 5,color = "black") +
  scale_x_continuous(breaks=seq(0, 100, by= 20)) +
  labs(x="Age", y="Frenquence") +
  ggtitle("Age Distribution")
```

Liver disease is high incidence between 35-60 years old. liver disese is related to age.

## **Is liver disease related to Age**
```{r}
liver_data %>% 
  ggplot(aes(liverdisease, Age))+
  geom_boxplot()+ geom_point()+
  ggtitle("Disease vs Age") +
  xlab("liverdisease") + ylab("Age")
``` 

 People with liver disease are older than those without disease.

## **Sex distribution in this data set**
```{r}
liver_data %>% group_by(Sex)%>%
  summarize(n=n())
``` 
More Male in this data set

## **liver disease summary by Sex**
```{r}
liver_data%>% ggplot(aes( liverdisease, fill= Sex)) +
  geom_bar()

liver_data%>% ggplot(aes( liverdisease, fill= Sex)) +
  geom_bar(position = position_dodge())
``` 

Based on the graph, male has high liver disease incident than female. 

## **Male with liver disease**
```{r}
liver_dataM<-liver_data %>% filter(Sex %in% 'Male')
Male_liverdiseace<-count(liver_dataM$Disease == "1") 
Male_liverdiseace
``` 
## **Female with liver disease**
```{r}
liver_dataF<-liver_data %>% filter(Sex %in% 'Female')
Female_liverdiseace <-count(liver_dataF$Disease == "1") 
Female_liverdiseace 
``` 

## **ratio of Male/Female in liver disease**
```{r}
Male_liverdiseace/Female_liverdiseace 
``` 


## **Male with no disease**
```{r}
Male_nodiseace<-count(liver_dataM$Disease == "0") 
Male_nodiseace
``` 

## **Female with liver disease**
```{r}
Female_nodiseace <-count(liver_dataF$Disease == "0") 
Female_nodiseace 
``` 
  
## **Male liver disease incident**  
```{r}
  Male_incident <- Male_liverdiseace/(Male_liverdiseace+Male_nodiseace)
  Male_incident
```   
## **Female liver incident**  
```{r}
  Female_incident <- Female_liverdiseace/(Male_liverdiseace+Male_nodiseace)
  Female_incident
```

## **liver disease incident ratio based on sex** 
```{r}
  Male_incident/Female_incident
``` 
  liver disease incident in Male is "r Male_incident/Female_incident" times higher than in Female. 
  
## **Tot_Bil express based on liverdisease**
```{r} 
  liver_data %>% 
    ggplot(aes(liverdisease, Tot_Bil))+
    geom_boxplot()+geom_point()+
    ggtitle("Disease vs Tot_Bil") +
    xlab("liverdisease") + ylab("Tot_Bil")
``` 

## **Tot_Bil expression based on liver disease (mean+/-2se)** 
```{r}
  liver_data %>% group_by(liverdisease) %>% 
    summarize(n=n(), avg=mean(Tot_Bil),sd=sd(Tot_Bil),se=sd(Tot_Bil)/sqrt(n()))%>%
    ggplot(aes(x=liverdisease, y=avg, ymin=avg-2*se, ymax=avg+2*se))+
    geom_point()+
    geom_errorbar(width=0.75, colour="black", alpha=0.75, size=0.25)+
    ggtitle("Tot_Bil")+
    xlab("liverdisease") + ylab("Tot_Bil")
```   

 Tot_Bil is higher in those have liver disease.
 
## **Dir_Bil expression based on liver disease---wrong** 
```{r}
  liver_data %>% 
    ggplot(aes(liverdisease, Dir_Bil))+
    geom_boxplot()+geom_point()+
    ggtitle("Disease vs Dir_Bil") +
    xlab("liverdisease") + ylab("Dir_Bil")
```
 
## **Dir_Bil expression based on liver disease (mean+/-2se)** 
```{r}
  liver_data %>% group_by(liverdisease) %>% 
    summarize(n=n(), avg=mean(Dir_Bil),sd=sd(Dir_Bil),se=sd(Dir_Bil)/sqrt(n()))%>%
    ggplot(aes(x=liverdisease, y=avg, ymin=avg-2*se, ymax=avg+2*se))+
    geom_point()+
    geom_errorbar(width=0.75, colour="black", alpha=0.75, size=0.25)+
    ggtitle("Dir_Bil")+
    xlab("liverdisease") + ylab("Dir_Bil")
```   

Dir_Bil is higher in  liver disease group 

## **Alkphos expression based on liver disease** 
```{r} 
  liver_data %>% 
    ggplot(aes(liverdisease,  Alkphos))+
    geom_boxplot()+geom_point()+
    ggtitle("Disease vs  Alkphos ") +
    xlab("liverdisease") + ylab(" Alkphos ")
```  

## **Alkphos  expression based on liver disease (mean+/-2se)** 
```{r}
  liver_data %>% group_by(liverdisease) %>% 
    summarize(n=n(), avg=mean( Alkphos),sd=sd( Alkphos ),se=sd( Alkphos )/sqrt(n()))%>%
    ggplot(aes(x=liverdisease, y=avg, ymin=avg-2*se, ymax=avg+2*se))+
    geom_point()+
    geom_errorbar(width=0.75, colour="black", alpha=0.75, size=0.25)+
    ggtitle(" Alkphos ")+
    xlab("liverdisease") + ylab(" Alkphos ")
```
  
Alkphos is higher in  liver disease group. 

## **Alamine expression based on liver disease** 
```{r}  
  liver_data %>% 
    ggplot(aes(liverdisease, Alamine ))+
    geom_boxplot()+geom_point()+
    ggtitle("Disease vs Alamine ") +
    xlab("liverdisease") + ylab(" Alamine ")
```

## **Alamine expression based on liver disease (mean+/-2se)**
```{r}
  liver_data %>% group_by(liverdisease) %>% 
    summarize(n=n(), avg=mean(Alamine ),sd=sd(Alamine),se=sd(Alamine)/sqrt(n()))%>%
    ggplot(aes(x=liverdisease, y=avg, ymin=avg-2*se, ymax=avg+2*se))+
    geom_point()+
    geom_errorbar(width=0.75, colour="black", alpha=0.75, size=0.25)+
    ggtitle("Alamine")+
    xlab("liverdisease") + ylab(" Alamine ")
```    
   
Alamine is higher in  liver disease group. 
   
## **Aspartate expression based on liver disease**
```{r}  
  liver_data %>% 
    ggplot(aes(liverdisease, Aspartate))+
    geom_boxplot()+geom_point()+
    ggtitle("Disease vs Aspartate") +
    xlab("liverdisease") + ylab(" Aspartate")
```  

## **Aspartate expression based on liver disease (mean+/-2se)**
```{r}  
  liver_data %>% group_by(liverdisease) %>% 
    summarize(n=n(), avg=mean(Aspartate),sd=sd(Aspartate),se=sd(Aspartate)/sqrt(n()))%>%
    ggplot(aes(x=liverdisease, y=avg, ymin=avg-2*se, ymax=avg+2*se))+
    geom_point()+
    geom_errorbar(width=0.75, colour="black", alpha=0.75, size=0.25)+
    ggtitle("Aspartate")+
    xlab("liverdisease") + ylab("Aspartate")
```   

Aspartate is higher in  liver disease group.   

## **Tot_Prot expression based on liver disease**
```{r} 
  liver_data %>% 
    ggplot(aes(liverdisease, Tot_Prot))+
    geom_boxplot()+geom_point()+
    ggtitle("Disease vs Tot_Prot") +
    xlab("liverdisease") + ylab("Tot_Prot")
```  

## **Tot_Prot expression based on liver disease (mean+/-2se)**
```{r}  
  liver_data %>% group_by(liverdisease) %>% 
    summarize(n=n(), avg=mean(Tot_Prot ),sd=sd(Tot_Prot),se=sd(Tot_Prot)/sqrt(n()))%>%
    ggplot(aes(x=liverdisease, y=avg, ymin=avg-2*se, ymax=avg+2*se))+
    geom_point()+
    geom_errorbar(width=0.75, colour="black", alpha=0.75, size=0.25)+
    ggtitle("Aspartate")+
    xlab("liverdisease") + ylab("Tot_Prot")
```  
Tot_Prot is lower in liver disease group. 

## **Albumin expression based on liver disease**
```{r}    
  liver_data %>% 
    ggplot(aes(liverdisease, Albumin))+
    geom_boxplot()+geom_point()+
    ggtitle("Disease vs Albumin") +
    xlab("liverdisease") + ylab("Albumin")
``` 

## **Albumin expression based on liver disease (mean+/-2se)**
```{r}   
  liver_data %>% group_by(liverdisease) %>% 
    summarize(n=n(), avg=mean(Albumin),sd=sd(Albumin),se=sd(Albumin)/sqrt(n()))%>%
    ggplot(aes(x=liverdisease, y=avg, ymin=avg-2*se, ymax=avg+2*se))+
    geom_point()+
    geom_errorbar(width=0.75, colour="black", alpha=0.75, size=0.25)+
    ggtitle("Albumin")+
    xlab("liverdisease") + ylab("Albumin")
```   

 Albumin is  lower in liver disease group. 
 
##**A_G_Ratio expression based on liver disease**
 ```{r}   
  liver_data %>% 
    ggplot(aes(liverdisease, A_G_Ratio))+
    geom_boxplot()+geom_point()+
    ggtitle("Disease vs A_G_Ratio") +
    xlab("liverdisease") + ylab("A_G_Ratio")
```  

## **A_G_Ratio expression based on liver disease (mean+/-2se)**
```{r}    
  liver_data %>% group_by(liverdisease) %>% 
    summarize(n=n(), avg=mean(A_G_Ratio),sd=sd(Albumin),se=sd(A_G_Ratio)/sqrt(n()))%>%
    ggplot(aes(x=liverdisease, y=avg, ymin=avg-2*se, ymax=avg+2*se))+
    geom_point()+
    geom_errorbar(width=0.75, colour="black", alpha=0.75, size=0.25)+
    ggtitle("A_G_Ratio")+
    xlab("liverdisease") + ylab("A_G_Ratio")
```   

A_G_Ratio is lower in liver disease group. 
 
 
# **Build the prediction system**

## **separete data set to train_set(70%) and test_set(30%)**
```{r}   
set.seed(1, sample.kind="Rounding") 
test_index1 <- createDataPartition(y = liver_data$Disease, times = 1, p = 0.3, list = FALSE)
train_set <- liver_data[-test_index1,]
test_set<- liver_data[test_index1,]
dim(train_set)
dim(test_set)
```

# ** build models**

Since "no disease" and  have "liver disease" is a "0" and "1" binary,   I think the Accuracy of each Model is more important to evaluate a model. But I will still calculate the rmse(residual mean square error).

# **model 1. liner Regression Model**

## **create liner Regression Model**
```{r}  
fit_lm <- lm(liverdisease ~ Age + Sex + Tot_Bil + Dir_Bil + Alkphos + Alamine + Aspartate + 
                 Tot_Prot + Albumin + A_G_Ratio, data = train_set)
fit_lm$coef
```

## **get a prediction based on liner Regression Model**
```{r}  
prediction_lm<- predict(fit_lm, test_set, type = "response")

Disease_lm <- ifelse(prediction_lm > 0.5, "1", "0") %>% factor
```

## **The Accuracy of liner Regression Model**

```{r}  
confusionMatrix(Disease_lm, test_set$liverdisease)$table
Accuracy_lm <- confusionMatrix(Disease_lm, test_set$liverdisease)$overall[["Accuracy"]]
Accuracy_lm
```

The accuracy of this liner Regression Model is `r Accuracy_lm`.

## **rmse of  liner Regression Model**
```{r}  
model_lm_rmse <- RMSE(prediction_lm, test_set$Disease)
model_lm_rmse
```

model_lm_rmse is `r model_lm_rmse`.

# **model 2. Logistic Regression Model**

## **create Logistic Regression Model**
```{r} 
fit_glm <- glm(liverdisease ~ Age + Sex + Tot_Bil + Dir_Bil + Alkphos + Alamine + Aspartate + 
                 Tot_Prot + Albumin + A_G_Ratio, data = train_set, family = binomial)
fit_glm$coef
```

## **get a prediction based on Logistic Regression Model**
```{r} 
prediction_glm<- predict(fit_glm, test_set, type = "response")

Disease_glm <- ifelse(prediction_glm > 0.5, "1", "0") %>% factor
```

## **The Accuracy of Logistic Regression Model**
```{r} 
confusionMatrix(Disease_glm, test_set$liverdisease)$table
Accuracy_glm <-confusionMatrix(Disease_glm, test_set$liverdisease)$overall[["Accuracy"]]
Accuracy_glm
```
The accuracy of this Logistic Regression Model is `r Accuracy_glm`.

## **rmse of  Logistic Regression Model**
```{r}  
model_glm_rmse <- RMSE(Accuracy_glm, test_set$Disease)
model_glm_rmse 
```

Model_glm_rmse is `r model_glm_rmse`.

# **model 3. knn Model**

## **find the best K for KNN model**
```{r} 
fit_knnK <- train(liverdisease ~ Age + Sex + Tot_Bil + Dir_Bil + Alkphos + Alamine + Aspartate + 
                Tot_Prot + Albumin + A_G_Ratio,   method = "knn", 
              tuneGrid = data.frame(k = seq(1, 15, 2)), 
              data = train_set)
ggplot(fit_knnK)

fit_knnK
```

When k=13, knn model get the highest Accuracy.

## **create Logistic Regression Model use k=13**
```{r} 
fit_knn <- knn3(liverdisease ~ Age + Sex + Tot_Bil + Dir_Bil + Alkphos + Alamine + Aspartate + 
                  Tot_Prot + Albumin + A_G_Ratio, data = train_set, k=13)
```

## **get a prediction based on knn Model**
```{r} 
prediction_knn <- predict(fit_knn, test_set, type = "class")
prediction_knn 
```

## **The Accuracy of knn Model**
```{r} 
confusionMatrix(data = prediction_knn, reference = test_set$liverdisease)$table
Accuracy_knn <- confusionMatrix(data = prediction_knn, reference = test_set$liverdisease)$overall["Accuracy"]

Accuracy_knn
```

The accuracy of this knn Model is "r Accuracy_knn".

## **rmse of knn Model**
```{r}  
model_knn_rmse <- RMSE(Accuracy_knn, test_set$Disease)
model_knn_rmse 

```

Model_knn_rmse  is `r model_knn_rmse`.

# **model 4. qda Model**

## **create qda Model**
```{r}
fit_qda <- train(liverdisease ~ Age + Sex + Tot_Bil + Dir_Bil + Alkphos + Alamine + Aspartate + 
                     Tot_Prot + Albumin + A_G_Ratio, data = train_set)
```

## **get a prediction based on qda Model**
```{r}
prediction_qda <- predict(fit_qda, test_set)
prediction_qda
```

## **compute accuracy**
```{r}
confusionMatrix(predict(fit_qda, test_set), test_set$liverdisease)$table
Accuracy_qda<-confusionMatrix(predict(fit_qda, test_set), test_set$liverdisease)$overall["Accuracy"]
Accuracy_qda
```

The accuracy of this qda Model is `r Accuracy_qda`.

## **rmse of qda Model**
```{r}
model_qda_rmse <- RMSE(Accuracy_qda, test_set$Disease)
model_qda_rmse 
```

Model_qda_rmse  is `r model_qda_rmse`.

# **model 5. lda Model**

## **create lda Model**
```{r}

fit_lda <- train(liverdisease ~ Age + Sex + Tot_Bil + Dir_Bil + Alkphos + Alamine + Aspartate + 
                   Tot_Prot + Albumin + A_G_Ratio, data = train_set)
```

## **get a prediction based on lda Model**
```{r}

prediction_lda <- predict(fit_lda, test_set)
prediction_lda

```

# **compute accuracy**
```{r}

confusionMatrix(predict(fit_qda, test_set), test_set$liverdisease)$table
Accuracy_lda <- confusionMatrix(predict(fit_lda, test_set), test_set$liverdisease)$overall["Accuracy"]
Accuracy_lda

```

The accuracy of this lda Model is `r Accuracy_lda`.

## **rmse of lda Model**
```{r}
model_lda_rmse <- RMSE(Accuracy_lda, test_set$Disease)
model_lda_rmse 
```

model_lda_rmse  is `r model_lda_rmse`.

# **model 6. Classification (Decision) Trees**

## **fit a classification tree and plot it**
```{r}
fit_tree <- train(liverdisease ~ Age + Sex + Tot_Bil + Dir_Bil + Alkphos + Alamine + Aspartate + 
                    Tot_Prot + Albumin + A_G_Ratio,
                     method = "rpart",
                     tuneGrid = data.frame(cp = seq(0.0, 0.1, len = 25)),
                     data = train_set)
plot(fit_tree)
```

## **get a prediction based on Decision Trees Model**
```{r}
prediction_Decision_Trees <- predict(fit_tree, test_set)
prediction_Decision_Trees
```

It get too much "1" (liver disease)

## **compute accuracy**
```{r}
confusionMatrix(predict(fit_tree, test_set), test_set$liverdisease)$table
Accuracy_Decision_Trees <- confusionMatrix(predict(fit_tree, test_set), test_set$liverdisease)$overall["Accuracy"]
Accuracy_Decision_Trees
```

Even the Accuracy_Decision_Trees is as highas `r Accuracy_Decision_Trees`, but the prediction is all "1"----liver disease, I think we should not trust this model. 

## **rmse of classification tree Model**
```{r}
model_Decision_Trees_rmse <- RMSE(Accuracy_Decision_Trees, test_set$Disease)
model_Decision_Trees_rmse 
```

# model 7. **randomForest Trees**
```{r}
library(randomForest)
```

## **create a randomForest tree model**
```{r}
fit_randomForest <- randomForest(liverdisease ~ Age + Sex + Tot_Bil + Dir_Bil + Alkphos + Alamine + Aspartate + 
                      Tot_Prot + Albumin + A_G_Ratio, data = train_set) 
```

## **get a prediction based on randomForest Trees**
```{r}
prediction_randomForest <- predict(fit_randomForest, test_set)
prediction_randomForest
```

## **compute accuracy**
```{r}
confusionMatrix(predict(fit_randomForest, test_set), test_set$liverdisease)$table
Accuracy_randomForest<- confusionMatrix(predict(fit_randomForest, test_set), test_set$liverdisease)$overall["Accuracy"]
Accuracy_randomForest
```

## **rmse of randomForest Model**
```{r}
model_randomForest_rmse <- RMSE(Accuracy_randomForest, test_set$Disease)
model_randomForest_rmse 
```

# **model 8.Regularization model**
## **create a test set that exclude "NA"**
```{r}   
test_set1 <- test_set %>% 
  semi_join(train_set, by = "Age") %>%
  semi_join(train_set, by = "Sex") %>%
  semi_join(train_set, by = "Tot_Bil")%>%
  semi_join(train_set, by = "Dir_Bil")%>%
  semi_join(train_set, by = "Alkphos")%>%
  semi_join(train_set, by = "Alamine")%>%
  semi_join(train_set, by = "Tot_Prot")%>%
  semi_join(train_set, by = "Aspartate")%>%
  semi_join(train_set, by = "Albumin")%>%
  semi_join(train_set, by = "A_G_Ratio")

dim(test_set1)
```

## **compute the average based on each predictor**
```{r} 
mu_Disease<- mean(train_set$Disease)
mu_Disease
```

### **average by Age**
```{r} 
Age_avgs <- train_set %>% 
  group_by(Age) %>% 
  summarize(b_Age = mean(Disease - mu_Disease))

Age_avgs
```

### **average by Sex**
```{r} 
Sex_avgs <- train_set %>% 
  left_join(Age_avgs, by='Age') %>%
  group_by(Sex) %>%
  summarize(b_Sex = mean(Disease - mu_Disease -b_Age ))

Sex_avgs
```

### **average by Tot_Bil** 
```{r} 
Tot_Bil_avgs <- train_set %>% 
  left_join(Age_avgs, by='Age') %>%
  left_join(Sex_avgs, by='Sex') %>%
  group_by(Tot_Bil) %>%
  summarize(b_Tot_Bil = mean(Disease - mu_Disease -b_Age-b_Sex))

Tot_Bil_avgs
```

### **average by Dir_Bil** 
```{r} 
Dir_Bil_avgs <- train_set %>% 
  left_join(Age_avgs, by='Age') %>%
  left_join(Sex_avgs, by='Sex') %>%
  left_join(Tot_Bil_avgs, by='Tot_Bil') %>%
  group_by(Dir_Bil) %>%
  summarize(b_Dir_Bil = mean(Disease - mu_Disease - b_Age - b_Sex -b_Tot_Bil ))

Dir_Bil_avgs
```

### **average by Alkphos** 
```{r} 
Alkphos_avgs <- train_set %>% 
  left_join(Age_avgs, by='Age') %>%
  left_join(Sex_avgs, by='Sex') %>%
  left_join(Tot_Bil_avgs, by='Tot_Bil') %>%
  left_join(Dir_Bil_avgs, by='Dir_Bil') %>%
  group_by(Alkphos) %>%
  summarize(b_Alkphos = mean(Disease - mu_Disease - b_Age - b_Sex -b_Tot_Bil - b_Dir_Bil ))

Alkphos_avgs
```

### **average by Alamine** 
```{r} 
Alamine_avgs <- train_set %>% 
  left_join(Age_avgs, by='Age') %>%
  left_join(Sex_avgs, by='Sex') %>%
  left_join(Tot_Bil_avgs, by='Tot_Bil') %>%
  left_join(Dir_Bil_avgs, by='Dir_Bil') %>%
  left_join(Alkphos_avgs, by='Alkphos') %>%
  group_by(Alamine) %>%
  summarize(b_Alamine = mean(Disease - mu_Disease - b_Age - b_Sex -b_Tot_Bil - b_Dir_Bil - b_Alkphos))

Alamine_avgs
```

### **average by Aspartate**
```{r} 
Aspartate_avgs <- train_set %>% 
  left_join(Age_avgs, by='Age') %>%
  left_join(Sex_avgs, by='Sex') %>%
  left_join(Tot_Bil_avgs, by='Tot_Bil') %>%
  left_join(Dir_Bil_avgs, by='Dir_Bil') %>%
  left_join(Alkphos_avgs, by='Alkphos') %>%
  left_join(Alamine_avgs, by='Alamine') %>%
  group_by(Aspartate) %>%
  summarize(b_Aspartate = mean(Disease - mu_Disease - b_Age - b_Sex -b_Tot_Bil - b_Dir_Bil - b_Alkphos - b_Alamine))

Aspartate_avgs
```

### **average by Tot_Prot**
```{r} 
Tot_Prot_avgs <- train_set %>% 
  left_join(Age_avgs, by='Age') %>%
  left_join(Sex_avgs, by='Sex') %>%
  left_join(Tot_Bil_avgs, by='Tot_Bil') %>%
  left_join(Dir_Bil_avgs, by='Dir_Bil') %>%
  left_join(Alkphos_avgs, by='Alkphos') %>%
  left_join(Alamine_avgs, by='Alamine') %>%
  left_join(Aspartate_avgs, by='Aspartate') %>%
  group_by(Tot_Prot) %>%
  summarize(b_Tot_Prot = mean(Disease - mu_Disease - b_Age - b_Sex -b_Tot_Bil - b_Dir_Bil - b_Alkphos - b_Alamine -b_Aspartate))

Tot_Prot_avgs
```

### **average by Albumin**
```{r} 
Albumin_avgs <- train_set %>% 
  left_join(Age_avgs, by='Age') %>%
  left_join(Sex_avgs, by='Sex') %>%
  left_join(Tot_Bil_avgs, by='Tot_Bil') %>%
  left_join(Dir_Bil_avgs, by='Dir_Bil') %>%
  left_join(Alkphos_avgs, by='Alkphos') %>%
  left_join(Alamine_avgs, by='Alamine') %>%
  left_join(Aspartate_avgs, by='Aspartate') %>%
  left_join(Tot_Prot_avgs, by='Tot_Prot') %>%
  group_by(Albumin) %>%
  summarize(b_Albumin = mean(Disease - mu_Disease - b_Age - b_Sex -b_Tot_Bil - b_Dir_Bil - b_Alkphos - b_Alamine - b_Aspartate - b_Tot_Prot))

Albumin_avgs
```

### **average by A_G_Ratio**
```{r} 
A_G_Ratio_avgs <- train_set %>% 
  left_join(Age_avgs, by='Age') %>%
  left_join(Sex_avgs, by='Sex') %>%
  left_join(Tot_Bil_avgs, by='Tot_Bil') %>%
  left_join(Dir_Bil_avgs, by='Dir_Bil') %>%
  left_join(Alkphos_avgs, by='Alkphos') %>%
  left_join(Alamine_avgs, by='Alamine') %>%
  left_join(Aspartate_avgs, by='Aspartate') %>%
  left_join(Tot_Prot_avgs, by='Tot_Prot') %>%
  left_join(Albumin_avgs, by='Albumin') %>%
  group_by(A_G_Ratio) %>%
  summarize(b_A_G_Ratio = mean(Disease - mu_Disease - b_Age - b_Sex -b_Tot_Bil - b_Dir_Bil - b_Alkphos - b_Alamine - b_Aspartate - b_Tot_Prot - b_Albumin))

A_G_Ratio_avgs
```

## **Data set for Regularization model**
```{r} 
Regularization_predicted_Disease <- test_set1 %>% 
  left_join(Age_avgs, by='Age') %>%
  left_join(Sex_avgs, by='Sex') %>%
  left_join(Tot_Bil_avgs, by='Tot_Bil') %>%
  left_join(Dir_Bil_avgs, by='Dir_Bil') %>%
  left_join(Alkphos_avgs, by='Alkphos') %>%
  left_join(Alamine_avgs, by='Alamine') %>%
  left_join(Aspartate_avgs, by='Aspartate') %>%
  left_join(Tot_Prot_avgs, by='Tot_Prot') %>%
  left_join(Albumin_avgs, by='Albumin') %>%
  left_join(A_G_Ratio_avgs, by='A_G_Ratio') %>%
  mutate(pred = mu_Disease + b_Age + b_Sex + b_Tot_Bil + b_Dir_Bil + b_Alkphos + b_Alamine + b_Aspartate + b_Tot_Prot + b_Albumin + b_A_G_Ratio) %>%
  .$pred

Regularization_predicted_Disease
```

## **transfer numeric to "0"--no disease, "1"--liver disease**
```{r} 
Regularization_predicted_Disease1 <- ifelse(Regularization_predicted_Disease> 0.5, "1", "0") %>% factor
Regularization_predicted_Disease1
```

## **compute accuracy**
```{r} 
confusionMatrix(Regularization_predicted_Disease1, test_set1$liverdisease)$table
Accuracy_Regularization <-confusionMatrix(Regularization_predicted_Disease1, test_set1$liverdisease)$overall[["Accuracy"]]
Accuracy_Regularization
```
The Accuracy of Regularization Model is not that high. 

## **rmse of Regularization_model**
```{r} 
Regularization_model_rmse <- RMSE(Regularization_predicted_Disease, test_set1$Disease)
Regularization_model_rmse
```

The Accuracy of this model is not as high as the previous one and the the rmse of this model is not as small as the previous one, so this model is not that good.


# **model 9. Tuning regularization model**
## **add Penalized least squares with different tuning**
I add tunning to penalty large estimates that come from small sample size.
```{r} 
lambdas_dl <- seq(0, 40, 0.5)

all_reg_lambda_dl_rmses <- sapply(lambdas_dl, function(l){
  
  mu_Disease<- mean(train_set$Disease)
  
  b_iAge <- train_set %>% 
    group_by(Age) %>% 
    summarize(b_iAge = sum(Disease - mu_Disease)/(n()+l))
  
  b_iSex <- train_set %>% 
    left_join(b_iAge, by='Age') %>%
    group_by(Sex) %>%
    summarize(b_iSex = mean(Disease - mu_Disease - b_iAge)/(n()+l))
  
  b_iTot_Bil <- train_set %>% 
    left_join(b_iAge, by='Age') %>%
    left_join(b_iSex, by='Sex') %>%
    group_by(Tot_Bil) %>%
    summarize(b_iTot_Bil = mean(Disease - mu_Disease - b_iAge - b_iSex)/(n()+l))
  
  b_iDir_Bil <- train_set %>% 
    left_join(b_iAge, by='Age') %>%
    left_join(b_iSex, by='Sex') %>%
    left_join(b_iTot_Bil, by='Tot_Bil') %>%
    group_by(Dir_Bil) %>%
    summarize(b_iDir_Bil = mean(Disease - mu_Disease - b_iAge - b_iSex - b_iTot_Bil)/(n()+l))
  
  b_iAlkphos <- train_set %>% 
    left_join(b_iAge, by='Age') %>%
    left_join(b_iSex, by='Sex') %>%
    left_join(b_iTot_Bil, by='Tot_Bil') %>%
    left_join(b_iDir_Bil, by='Dir_Bil') %>%
    group_by(Alkphos) %>%
    summarize(b_iAlkphos = mean(Disease - mu_Disease - b_iAge - b_iSex - b_iTot_Bil- b_iDir_Bil)/(n()+l))
  
  b_iAlamine <- train_set %>% 
    left_join(b_iAge, by='Age') %>%
    left_join(b_iSex, by='Sex') %>%
    left_join(b_iTot_Bil, by='Tot_Bil') %>%
    left_join(b_iDir_Bil, by='Dir_Bil') %>%
    left_join(b_iAlkphos, by='Alkphos') %>%
    group_by(Alamine) %>%
    summarize(b_iAlamine = mean(Disease - mu_Disease - b_iAge - b_iSex - b_iTot_Bil- b_iDir_Bil - b_iAlkphos)/(n()+l))
  
  b_iAspartate <- train_set %>% 
    left_join(b_iAge, by='Age') %>%
    left_join(b_iSex, by='Sex') %>%
    left_join(b_iTot_Bil, by='Tot_Bil') %>%
    left_join(b_iDir_Bil, by='Dir_Bil') %>%
    left_join(b_iAlkphos, by='Alkphos') %>%
    left_join(b_iAlamine, by='Alamine') %>%
    group_by(Aspartate) %>%
    summarize(b_iAspartate = mean(Disease - mu_Disease - b_iAge - b_iSex - b_iTot_Bil- b_iDir_Bil - b_iAlkphos - b_iAlamine)/(n()+l))
  
  
  b_iTot_Prot <- train_set %>% 
    left_join(b_iAge, by='Age') %>%
    left_join(b_iSex, by='Sex') %>%
    left_join(b_iTot_Bil, by='Tot_Bil') %>%
    left_join(b_iDir_Bil, by='Dir_Bil') %>%
    left_join(b_iAlkphos, by='Alkphos') %>%
    left_join(b_iAlamine, by='Alamine') %>%
    left_join(b_iAspartate, by='Aspartate') %>%
    group_by(Tot_Prot) %>%
    summarize(b_iTot_Prot = mean(Disease - mu_Disease - b_iAge - b_iSex - b_iTot_Bil- b_iDir_Bil - b_iAlkphos - b_iAlamine - b_iAspartate)/(n()+l))
  
  
  b_iAlbumin <- train_set %>% 
    left_join(b_iAge, by='Age') %>%
    left_join(b_iSex, by='Sex') %>%
    left_join(b_iTot_Bil, by='Tot_Bil') %>%
    left_join(b_iDir_Bil, by='Dir_Bil') %>%
    left_join(b_iAlkphos, by='Alkphos') %>%
    left_join(b_iAlamine, by='Alamine') %>%
    left_join(b_iAspartate, by='Aspartate') %>%
    left_join(b_iTot_Prot, by='Tot_Prot') %>%
    group_by(Albumin) %>%
    summarize(b_iAlbumin = mean(Disease - mu_Disease - b_iAge - b_iSex - b_iTot_Bil- b_iDir_Bil - b_iAlkphos - b_iAlamine - b_iAspartate - b_iTot_Prot)/(n()+l))
  
  
  b_iA_G_Ratio <- train_set %>% 
    left_join(b_iAge, by='Age') %>%
    left_join(b_iSex, by='Sex') %>%
    left_join(b_iTot_Bil, by='Tot_Bil') %>%
    left_join(b_iDir_Bil, by='Dir_Bil') %>%
    left_join(b_iAlkphos, by='Alkphos') %>%
    left_join(b_iAlamine, by='Alamine') %>%
    left_join(b_iAspartate, by='Aspartate') %>%
    left_join(b_iTot_Prot, by='Tot_Prot') %>%
    left_join(b_iAlbumin, by='Albumin') %>%
    group_by(A_G_Ratio) %>%
    summarize(b_iA_G_Ratio = mean(Disease - mu_Disease - b_iAge - b_iSex - b_iTot_Bil- b_iDir_Bil - b_iAlkphos - b_iAlamine - b_iAspartate - b_iTot_Prot - b_iAlbumin)/(n()+l))
  
  Regularization_d_predicted_Disease <- test_set1 %>% 
    left_join(b_iAge, by='Age') %>%
    left_join(b_iSex, by='Sex') %>%
    left_join(b_iTot_Bil, by='Tot_Bil') %>%
    left_join(b_iDir_Bil, by='Dir_Bil') %>%
    left_join(b_iAlkphos, by='Alkphos') %>%
    left_join(b_iAlamine, by='Alamine') %>%
    left_join(b_iAspartate, by='Aspartate') %>%
    left_join(b_iTot_Prot, by='Tot_Prot') %>%
    left_join(b_iAlbumin, by='Albumin') %>%
    left_join(b_iA_G_Ratio, by='A_G_Ratio') %>%
    mutate(pred = mu_Disease + b_iAge + b_iSex + b_iTot_Bil+ b_iDir_Bil + b_iAlkphos + b_iAlamine + b_iAspartate + b_iTot_Prot + b_iAlbumin + b_iA_G_Ratio) %>%
    pull(pred)
  
  Regularization_d_predicted_Disease1 <- ifelse(Regularization_d_predicted_Disease> 0.5, "1", "0") %>% factor

  
  return(RMSE(Regularization_d_predicted_Disease, test_set1$Disease))
})

all_reg_lambda_dl_rmses

```

## **summary λ with corresponding rmse and find the lambda that get the smallest rmse**
```{r} 
qplot(lambdas_dl, all_reg_lambda_dl_rmses) 

lambda_rmse <- lambdas_dl [which.min(all_reg_lambda_dl_rmses)]
lambda_rmse

all_reg_lambda_dl_rmses[which.min(all_reg_lambda_dl_rmses)]
```

When lambda=`r lambda_rmse`, we got the smallest rmse  `r which.min(all_reg_lambda_dl_rmses)`. It seems add Penalized least squares tuning not improve rsme. 

## **prediction Accuracy by add Penalized least squares with different tuning**
```{r} 
lambdas_dl <- seq(0, 40, 0.5)

all_reg_lambda_dl_Accuracy <- sapply(lambdas_dl, function(k){
  
  mu_Disease<- mean(train_set$Disease)
  
  b_iAge <- train_set %>% 
    group_by(Age) %>% 
    summarize(b_iAge = sum(Disease - mu_Disease)/(n()+k))
  
  b_iSex <- train_set %>% 
    left_join(b_iAge, by='Age') %>%
    group_by(Sex) %>%
    summarize(b_iSex = mean(Disease - mu_Disease - b_iAge)/(n()+k))
  
  b_iTot_Bil <- train_set %>% 
    left_join(b_iAge, by='Age') %>%
    left_join(b_iSex, by='Sex') %>%
    group_by(Tot_Bil) %>%
    summarize(b_iTot_Bil = mean(Disease - mu_Disease - b_iAge - b_iSex)/(n()+k))
  
  b_iDir_Bil <- train_set %>% 
    left_join(b_iAge, by='Age') %>%
    left_join(b_iSex, by='Sex') %>%
    left_join(b_iTot_Bil, by='Tot_Bil') %>%
    group_by(Dir_Bil) %>%
    summarize(b_iDir_Bil = mean(Disease - mu_Disease - b_iAge - b_iSex - b_iTot_Bil)/(n()+k))
  
  b_iAlkphos <- train_set %>% 
    left_join(b_iAge, by='Age') %>%
    left_join(b_iSex, by='Sex') %>%
    left_join(b_iTot_Bil, by='Tot_Bil') %>%
    left_join(b_iDir_Bil, by='Dir_Bil') %>%
    group_by(Alkphos) %>%
    summarize(b_iAlkphos = mean(Disease - mu_Disease - b_iAge - b_iSex - b_iTot_Bil- b_iDir_Bil)/(n()+k))
  
  b_iAlamine <- train_set %>% 
    left_join(b_iAge, by='Age') %>%
    left_join(b_iSex, by='Sex') %>%
    left_join(b_iTot_Bil, by='Tot_Bil') %>%
    left_join(b_iDir_Bil, by='Dir_Bil') %>%
    left_join(b_iAlkphos, by='Alkphos') %>%
    group_by(Alamine) %>%
    summarize(b_iAlamine = mean(Disease - mu_Disease - b_iAge - b_iSex - b_iTot_Bil- b_iDir_Bil - b_iAlkphos)/(n()+k))
  
  b_iAspartate <- train_set %>% 
    left_join(b_iAge, by='Age') %>%
    left_join(b_iSex, by='Sex') %>%
    left_join(b_iTot_Bil, by='Tot_Bil') %>%
    left_join(b_iDir_Bil, by='Dir_Bil') %>%
    left_join(b_iAlkphos, by='Alkphos') %>%
    left_join(b_iAlamine, by='Alamine') %>%
    group_by(Aspartate) %>%
    summarize(b_iAspartate = mean(Disease - mu_Disease - b_iAge - b_iSex - b_iTot_Bil- b_iDir_Bil - b_iAlkphos - b_iAlamine)/(n()+k))
  
  
  b_iTot_Prot <- train_set %>% 
    left_join(b_iAge, by='Age') %>%
    left_join(b_iSex, by='Sex') %>%
    left_join(b_iTot_Bil, by='Tot_Bil') %>%
    left_join(b_iDir_Bil, by='Dir_Bil') %>%
    left_join(b_iAlkphos, by='Alkphos') %>%
    left_join(b_iAlamine, by='Alamine') %>%
    left_join(b_iAspartate, by='Aspartate') %>%
    group_by(Tot_Prot) %>%
    summarize(b_iTot_Prot = mean(Disease - mu_Disease - b_iAge - b_iSex - b_iTot_Bil- b_iDir_Bil - b_iAlkphos - b_iAlamine - b_iAspartate)/(n()+k))
  
  
  b_iAlbumin <- train_set %>% 
    left_join(b_iAge, by='Age') %>%
    left_join(b_iSex, by='Sex') %>%
    left_join(b_iTot_Bil, by='Tot_Bil') %>%
    left_join(b_iDir_Bil, by='Dir_Bil') %>%
    left_join(b_iAlkphos, by='Alkphos') %>%
    left_join(b_iAlamine, by='Alamine') %>%
    left_join(b_iAspartate, by='Aspartate') %>%
    left_join(b_iTot_Prot, by='Tot_Prot') %>%
    group_by(Albumin) %>%
    summarize(b_iAlbumin = mean(Disease - mu_Disease - b_iAge - b_iSex - b_iTot_Bil- b_iDir_Bil - b_iAlkphos - b_iAlamine - b_iAspartate - b_iTot_Prot)/(n()+k))
  
  
  b_iA_G_Ratio <- train_set %>% 
    left_join(b_iAge, by='Age') %>%
    left_join(b_iSex, by='Sex') %>%
    left_join(b_iTot_Bil, by='Tot_Bil') %>%
    left_join(b_iDir_Bil, by='Dir_Bil') %>%
    left_join(b_iAlkphos, by='Alkphos') %>%
    left_join(b_iAlamine, by='Alamine') %>%
    left_join(b_iAspartate, by='Aspartate') %>%
    left_join(b_iTot_Prot, by='Tot_Prot') %>%
    left_join(b_iAlbumin, by='Albumin') %>%
    group_by(A_G_Ratio) %>%
    summarize(b_iA_G_Ratio = mean(Disease - mu_Disease - b_iAge - b_iSex - b_iTot_Bil- b_iDir_Bil - b_iAlkphos - b_iAlamine - b_iAspartate - b_iTot_Prot - b_iAlbumin)/(n()+k))
  
  Regularization_d_predicted_Disease <- test_set1 %>% 
    left_join(b_iAge, by='Age') %>%
    left_join(b_iSex, by='Sex') %>%
    left_join(b_iTot_Bil, by='Tot_Bil') %>%
    left_join(b_iDir_Bil, by='Dir_Bil') %>%
    left_join(b_iAlkphos, by='Alkphos') %>%
    left_join(b_iAlamine, by='Alamine') %>%
    left_join(b_iAspartate, by='Aspartate') %>%
    left_join(b_iTot_Prot, by='Tot_Prot') %>%
    left_join(b_iAlbumin, by='Albumin') %>%
    left_join(b_iA_G_Ratio, by='A_G_Ratio') %>%
    mutate(pred = mu_Disease + b_iAge + b_iSex + b_iTot_Bil+ b_iDir_Bil + b_iAlkphos + b_iAlamine + b_iAspartate + b_iTot_Prot + b_iAlbumin + b_iA_G_Ratio) %>%
    pull(pred)
  
  Regularization_d_predicted_Disease1 <- ifelse(Regularization_d_predicted_Disease> 0.5, "1", "0") %>% factor
  
  
  return(confusionMatrix(Regularization_d_predicted_Disease1, test_set1$liverdisease)$overall[["Accuracy"]])
  
})
all_reg_lambda_dl_Accuracy 

```

## **the Accuracy of Tuning_regularization_model**
```{r} 
lambdas_dl [which.min(all_reg_lambda_dl_rmses)]
all_reg_lambda_dl_Accuracy[which.min(all_reg_lambda_dl_rmses)]

```

## **the rmse of Tuning_regularization_model**
```{r} 
lambdas_dl [which.max(all_reg_lambda_dl_Accuracy)]
all_reg_lambda_dl_rmses[which.max(all_reg_lambda_dl_Accuracy)]
```


## **the highest accuracy of Tuning_regularization_model**
```{r} 
the_highest_acccuracy_of_Tuning_regularization_model<-  all_reg_lambda_dl_Accuracy [which.max(all_reg_lambda_dl_Accuracy)]
the_highest_acccuracy_of_Tuning_regularization_model
```

## **the smallest rmse of Tuning_regularization_model**
```{r} 
the_smallest_rmse_of_all_Tuning_regularization_model <- all_reg_lambda_dl_rmses[which.min(all_reg_lambda_dl_rmses)]
the_smallest_rmse_of_all_Tuning_regularization_model
```

Based on the highest accuracy and the smallest rsme, this model does not looks like a good one.


# **III.Result**

I try 9 models, summary as the following: 
## ** summary of accuracy and rmse **
```{r}

asscracy_rmse_summary<-data_frame ( Method = c("liner Regression Model", "Logistic Regression Model","knn Model", "qda Model","lda Model", "Decision Trees", "random Forest Trees", "Regularization model", "Tuning regularization model"), Accuracy = c(Accuracy_lm, Accuracy_glm, Accuracy_knn, Accuracy_qda, Accuracy_lda, Accuracy_Decision_Trees, Accuracy_randomForest, Accuracy_Regularization, the_highest_acccuracy_of_Tuning_regularization_model), rmse = c( model_lm_rmse, model_glm_rmse, model_knn_rmse, model_qda_rmse, model_lda_rmse, model_Decision_Trees_rmse, model_randomForest_rmse, Regularization_model_rmse, the_smallest_rmse_of_all_Tuning_regularization_model))

asscracy_rmse_summary

```

The 2 highest accuracy models ---liner Regression Model and Decision Trees give all prediction  "1"--liver disease, this is not true. So I will exclude these 2 models. When running the Regularization model and Tuning regularization model, I remove "NA"from the data set, this make the data size too small, which is only  `r  dim(test_set1)`, maybe this is the reason the accuracy of Regularization model is not that high, which is `r Accuracy_Regularization`, and rmse `r Regularization_model_rmse` is larger compare to other model.  In Tuning regularization model model, the highest accuracy and the smallest rmse are not with the same lambda. Tuning regularization model get accuracy of `r the_highest_acccuracy_of_Tuning_regularization_model` which is not the highest of all the 9 model and the corresponding rmse is `r all_reg_lambda_dl_rmses[which.max(all_reg_lambda_dl_Accuracy)]`. 
Thinking of both accuracy and rmsem,  qda Model get the highest accuracy `r Accuracy_qda` and the smallest rmse `r model_qda_rmse`, the prediction table looks reasonable too. So I think the qda model is the best prediction model for this liver disease data set. 


# **IV. Conclusion**

This project is to build a model that can predict a liver disease status based on features of "Age", "Sex", "Tot_Bil", "Dir_Bil", "Alkphos", "Alamine", 
"Aspartate", "Tot_Prot", "Albumin", "A_G_Ratio".  After review the data set, I find all these features are related to disease status. So we use all these features in building models. 9 models are created----liner Regression Model, Logistic Regression Model, knn Model, qda Model, lda Model, Decision Trees, random Forest Trees, Regularization model, Tuning regularization model, the qda Model is the best one with a accuracy of `r Accuracy_qda` and a rmse of `r model_qda_rmse`.  Building of a disease prediction model will create a new method for disease diagnosis in public health and clinic. 

Limitation: These data set size is not large enough to get good prdiction, espically when I exculded missing and "NA" value.  

In the future, as more samples are added to this data set, we can try again. or maybe try some more model, for example, PCA. 